---
title: "Annotate and Summary Stats of Breakpoint Evidence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Annotate and Summary Stats of Breakpoint Evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

#Set-up

```{r setup}
knitr::opts_knit$set(
  root.dir = file.path(PROJHOME,"2020.09.15_RNAseq_Fusion_Breakpoints"))

library(fusBreakpoint)
library(magrittr)
library(dplyr)
library(tibble)
library(stringr)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts=list(width.cutoff=50),
  tidy=TRUE, 
  fig.align='center',
  fig.width = 10, 
  fig.height = 10
)
```


#Define Functions

```{r}
#Needs to be build into the package
source(file.path(SCRIPTS,"RNAseq_Analysis/fusBreakpoint/R/survplot_functions.R.bak"))
```

#ClinData

```{r}
merged <- read.csv(file.path(CDE,"Merged/TARGET_AML_0531_1031_merged_CDEs_9.18.20.csv"))

merged <- merged %>%
  filter(!is.na(USI), USI != "Unknown") %>%
  set_rownames(.$USI)

inelig <- c(773920,775026, 786948,799528)
ineligables <- merged %>% 
  filter(Eligibility_Comments == "remove" | Reg. %in% inelig) %>% 
  pull(USI)

dim(merged)
```

```{r}
IDmap <- read.csv("References/St.Jude_TARGET_CICERO_Sample_IDmap.csv")

head(IDmap)
```
 
```{r}
SJCBF <- read.csv("References/SJ_CBF_AML_Clinical_Data.csv") %>% 
  mutate(Sample=str_split_fixed(ID, "\\/", n=2)[,1]) %>% 
  mutate(subject_name=str_split_fixed(Sample,"_", n=2)[,1]) %>%
  select(Sample,subject_name, ID, everything())

# head(SJCBF)
dim(SJCBF)
```


#Read in File Manifests 

```{r}
polyA_RNAseq_files <- read.csv("BAM_Manifests/TARGET_AML_polyA_RNAseq_Bam_Manifest_10.02.20.csv", 
                               row.names = 1) 


dim(polyA_RNAseq_files)
table(duplicated(polyA_RNAseq_files$Sample)) #FALSE
```

```{r}
RBD_RNAseq_files <- read.csv("BAM_Manifests/TARGET_AML_Ribodepleted_RNAseq_Bam_Manifest_10.02.20.csv",row.names = 1)

dim(RBD_RNAseq_files)
table(duplicated(RBD_RNAseq_files$Sample)) #FALSE
```

```{r}
SJ_RNAseq_files <- read.csv("BAM_Manifests/St.Jude_AML_ALL_RNAseq_Bam_Manifest_10.09.20.csv")

dim(SJ_RNAseq_files)
table(duplicated(SJ_RNAseq_files$Sample)) #FALSE
```


#Define the Results files 

```{r}
outdir <- file.path(SCRATCH,'jlsmith3/Fusion_Breakpoints')
outdir
```

```{r}
results_files <- dir(file.path(outdir), 
                 pattern = "results_[01].RDS",
                 recursive = T, 
                 full.names = T)

Names <- str_split_fixed(results_files, "\\/", n=9)[,8] 
Names <- str_split_fixed(Names,"_", n=6)[,3:4]
Names <- paste(as.character(Names[,1]),
               as.character(Names[,2]), sep=".")
names(results_files) <- Names

length(results_files) # 62
head(results_files)
table(names(results_files))
```

#Save TARGET AML Results

```{r}
TARGET_results <- grep("polyA_|RBD_", results_files, value = T)

table(names(TARGET_results))
length(TARGET_results)
```

```{r eval=FALSE}
cat_results <- lapply(seq(2,length(TARGET_results), by=2), 
                      function(i){

                #select the results files  (in two parts)
                rds0 <- TARGET_results[i-1]
                rds1 <- TARGET_results[i]
                # print(c(rds0,rds1))
                
                #define output filename and destination
                fname <- str_split_fixed(rds0,"\\/", n=9)[,8] %>% 
                  gsub("_rslurm_", "", .) %>% 
                  paste0("TARGET_AML_",., "_breakpoint_evidence.csv")
                
                fusion <- unique(c(names(rds0),names(rds1)))
                print(fusion)
                destination <- paste0(getwd(), "/", fusion, "_Results")
                suppressWarnings(dir.create(destination,recursive = T))
                
                #read in the results
                res <- purrr::map_dfr(c(rds0, rds1), function(x) bind_rows(readRDS(x)))
                
                # RNAseq batch annotations
                if(grepl("polyA", rds0)){
                  res <- res %>%
                      left_join(., select(polyA_RNAseq_files, -filename, -filepath),
                              by="Sample") 
                }else{
                  res <- res %>%
                      left_join(., select(RBD_RNAseq_files, -filename, -filepath),
                                by="Sample") 
                }
                
                #merge in clinical annotations
                res <- res %>%
                     left_join(., select(merged, USI,
                                            matches("OS\\.[tei]|EFS\\.[tei]|Event",ignore.case=FALSE)),
                                  by="USI") %>%
                      mutate_at(vars(Primary.Fusion), ~ifelse(grepl("NBM|PB", Group), Group, .)) %>% 
                      select(Sample,USI,Protocol,
                               Primary.Fusion,Additional.Fusions.CNV,
                               Theoretical_Exon_Junctions,
                              Num_Junction_Breakpoint_Reads,
                              Num_RNAseq_Reads_Searched,
                              Lib_Prep:AML_Subtype, everything(),
                             -Reg., -Final_Patient_ID, -PATIENT_ID_Original)
                
                print(paste(destination, fname, sep="/"))
                write.csv(res,paste(destination, fname, sep="/"), row.names = FALSE)
              
              })
```


#Analysis of Breakpoint Evidence


##CBFB-MYH11

```{r}
cbfb_mut20_RBD <- read.csv("CBFB.MYH11_Results/TARGET_AML_CBFB_MYH11_RBD_RNAseq_mut20_breakpoint_evidence.csv") %>% 
  mutate(Exons=gsub(".+Exons([0-9]_[0-9]{1,2})_.+", "\\1", Theoretical_Exon_Junctions)) %>%
  dplyr::select(Sample:Additional.Fusions.CNV, Exons,Num_Junction_Breakpoint_Reads, everything())

head(cbfb_mut20_RBD)
length(unique(cbfb_mut20_RBD$Sample)) #1617
```

```{r}
cbfb_mut20_polyA <- read.csv("CBFB.MYH11_Results/TARGET_AML_CBFB_MYH11_polyA_RNAseq_mut20_breakpoint_evidence.csv") %>% 
  mutate(Exons=gsub(".+Exons([0-9]_[0-9]{1,2})_.+", "\\1", Theoretical_Exon_Junctions)) %>%
  dplyr::select(Sample:Additional.Fusions.CNV, Exons,Num_Junction_Breakpoint_Reads, everything())

head(cbfb_mut20_polyA)
length(unique(cbfb_mut20_polyA$Sample)) # 508
```


```{r}
cbfb_all <- cbfb_mut20_RBD %>% 
  bind_rows(., cbfb_mut20_polyA) %>% 
  # filter(!USI %in% ineligables) %>%
  
  group_by(Sample,Lib_Prep) %>%
  mutate(CBFB_MYH11_True_Positive=case_when(
   (grepl("CBFB-MYH11", Primary.Fusion) | grepl("CBFB-MYH11", Additional.Fusions.CNV)) & any(Num_Junction_Breakpoint_Reads > 0) ~ "TP",
   !(grepl("CBFB-MYH11", Primary.Fusion) | grepl("CBFB-MYH11", Additional.Fusions.CNV)) & any(Num_Junction_Breakpoint_Reads > 0) ~ "FP", 
   (grepl("CBFB-MYH11", Primary.Fusion) | grepl("CBFB-MYH11", Additional.Fusions.CNV)) & any(Num_Junction_Breakpoint_Reads == 0) ~ "FN", 
   !(grepl("CBFB-MYH11", Primary.Fusion) | grepl("CBFB-MYH11", Additional.Fusions.CNV)) & any(Num_Junction_Breakpoint_Reads == 0) ~ "TN")) %>% 
  
  ungroup() %>%
  dplyr::select(Sample:Additional.Fusions.CNV,CBFB_MYH11_True_Positive, everything())


# dim(cbfb_all) #10625    34
# length(unique(cbfb_all$Sample)) #1967 (1961 without ineligables)
head(cbfb_all)
```


## Check with the original 

```{r}
cbfb_res_orig <- xlsx::read.xlsx("CBFB.MYH11_Results/Theoretical_data/TARGET_0531_1031.CBFB-MYH11.FusionBreakpoint.Unmasked.xlsx", sheetIndex = 1)

# head(cbfb_res_orig)
dim(cbfb_res_orig)#152   4
length(unique(cbfb_res_orig$USI))
```

```{r}
check <- cbfb_res_orig  %>% 
  inner_join(., cbfb_all,
             by="USI") %>%
  select(Sample:Exons,CBFB_MYH11_True_Positive,
         Fusion_Breakpoint_Unmasked, Fusion_Breakpoint, everything()) %>%
  
  group_by(Sample, Lib_Prep) %>%
  mutate(Keep=case_when(
    Num_Junction_Breakpoint_Reads > 0 ~ TRUE, 
    Num_Junction_Breakpoint_Reads == 0 & CBFB_MYH11_True_Positive == "FN" ~ !duplicated(Sample),
    TRUE ~ FALSE)) %>%
  ungroup() %>%
  # filter(Num_Junction_Breakpoint_Reads > 0) %>%
  
  # mutate(Different=ifelse(Exons==Fusion_Breakpoint_Unmasked, "No", "Yes")) %>%
  # arrange(desc(Fusion_Breakpoint), Fusion_Breakpoint_Unmasked, USI) %>%
  select(Keep, Sample:Additional.Fusions.CNV,Lib_Prep,
         # Different,
         everything())

check




# table(check$Different, useNA='ifany')
# table(duplicated(check$USI))
# write.csv(check,"CBFB_MYH11_Check.csv", row.names = FALSE)
```


```{r}
toFill <- cbfb_all %>% 
  filter(CBFB_MYH11_True_Positive=="TP", Num_Junction_Breakpoint_Reads >0) %>% 
  filter(! USI %in% cbfb_res_orig$USI) %>% 
  select(Sample,USI, Exons)

# toFill
dim(toFill)
# write.csv(toFill,"CBFB.MYH11_Results/TARGET_AML_CBFB-MYH11_RNAseq_mut20_bySeq_Search.csv", row.names = FALSE)
```



## Specificty and Sensitivity


```{r}
cbfb_all %>% 
  select(Sample,CBFB_MYH11_True_Positive) %>% 
  unique() %>% 
  group_by(CBFB_MYH11_True_Positive) %>% 
  summarise(N=n()) %>% 
  ungroup()
```

```{r}
FN_all <- filter(cbfb_all, CBFB_MYH11_True_Positive == "FN") %>% 
  arrange(Protocol) %>% 
  dplyr::select(Sample,USI,Protocol, Lib_Prep,
                Primary.Fusion, Additional.Fusions.CNV, CBFB_MYH11_True_Positive)  %>% 
  unique()

FN_all #13 False negatives (not so great)

# table(FN_all$USI %in% ineligables)
# table(FN_all$USI %in% cbfb_res_orig$USI)
```

TARGET.20.PARAHF.09A.01R	#Low depth
TARGET.20.PASGKA.09A.01R # Low depth 
TARGET.20.PASVVM.09A.01R #low depth

TARGET.20.PAURKG.09A.01R #RBD
TARGET.20.PAWHTL.03A.01R #RBD 
TARGET.20.PAWNIK.09A.01R #RBD



# ST.Jude Data

```{r}
#Need STAR fusion data in order to see how many Inv.16 were detected. 
star.fmt.primary <- read.csv(file.path("St.Jude_AML_ALL_STAR_Fusion_reformatted_FilteredForNBM_PrimaryFusions_10.12.20.csv"),
                             row.names=FALSE)
```


```{r}
SJ_results <- grep("_SJ_", results_files, value = T)
# SJresults
table(names(SJ_results))
```

```{r eval=FALSE}
cat_results <- lapply(seq(2,length(SJ_results), by=2), 
                      function(i){

                #select the results files  (in two parts)
                rds0 <- SJ_results[i-1]
                rds1 <- SJ_results[i]
                print(c(rds0,rds1))
                
                #define output filename and destination
                fname <- str_split_fixed(rds0,"\\/", n=9)[,8] %>% 
                  gsub("_rslurm_", "", .) %>% 
                  paste0("St.Jude_AML_ALL_",., "_breakpoint_evidence.csv")
                
                fusion <- unique(c(names(rds0),names(rds1)))
                print(fusion)
                destination <- paste0(getwd(), "/", fusion, "_Results")
                # suppressWarnings(dir.create(destination,recursive = T))
                
                # #read in the results
                # res <- purrr::map_dfr(c(rds0, rds1), function(x) bind_rows(readRDS(x)))
                # 
                # # RNAseq batch annotations
                # res <- res %>%
                #       left_join(., select(SJ_RNAseq_files, -filename, -filepath),
                #               by="Sample") 
                # 
                
                print(paste(destination, fname, sep="/"))
                # write.csv(res,paste(destination, fname, sep="/"), row.names = FALSE)
              
})
```




#Session Information

```{r}
sessionInfo()
```



