---
title: "Multiple_Samples_with_RSlurm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multiple_Samples_with_RSlurm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Set-up

```{r setup, message=FALSE}
knitr::opts_knit$set(
  root.dir = file.path(PROJHOME,"2020.09.15_RNAseq_Fusion_Breakpoints"))

library(fusBreakpoint)
library(magrittr)
library(dplyr)
library(tibble)
library(stringr)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts=list(width.cutoff=50),
  tidy=TRUE, 
  fig.align='center',
  fig.width = 10, 
  fig.height = 10
)
```


# Define Functions

```{r}
#Function to wrap up all the RSlurm submissions for the different fusion groups and fusion sequences. 
submit_slurm_per_fusion <- function(fusion.category, file_manifest, chromLengths, cohort=""){
  # The fusion group to investigate
  group <- fusion.category
  breakpoint <- common_breakpoints[group,"breakpoint"]
  print(group)
  print(breakpoint)
  
  #Filter the sequence data for the fusion to investigate
  fusion_theoretical_input <- filter(theoretical_data.collapsed, Fusion.Category == group)
  cicero.contig_input <- filter(cicero.contig, Fusion.Category ==  group)
  TA.contigs_input <- filter(TA.contigs, Fusion.Category == group)
  
  # I forgot seqs need to be the same width for pdict objects. hopefully will change soon.
  try(
    mut20 <- create_custom_pdict(theoretical_seqs_df = filter(fusion_theoretical_input,
                                                              grepl("Mutseq20", Name))), silent=T)
  try(
    mut25 <- create_custom_pdict(theoretical_seqs_df = filter(fusion_theoretical_input,
                                                            !grepl("Mutseq20", Name))), silent = T)
  
  for (n in 1:2){
  
    mutseq <- c("mut20","mut25")[n]
    try(pdict_obj <- get(mutseq), silent=T)
    
    if(exists("pdict_obj")){ #some groups do not have both types of data.
      samps <- as.list(pull(file_manifest, Sample))
      jobname <- paste(gsub("-","_",group),
                       cohort,mutseq, sep="_")
      message(paste0("submitting: ",jobname))
      
      sjob <- rslurm::slurm_map(samps,
                      examine_breakpoint_evidence,
                      breakpoint = breakpoint,
                      file_manifest = file_manifest,
                      Seqlens = chromLengths,
                      fusion_theoretical_pdict = pdict_obj,
                      cicero.contig = cicero.contig_input,
                      TA.contigs = TA.contigs_input,
                      slurm_options=sopt,
                      jobname=jobname,
                      submit = T)
      rm(pdict_obj)
    }
    rm(mutseq)
  }
}
```

# ClinData

```{r}
merged <- read.csv(file.path(CDE,"Merged/TARGET_AML_0531_1031_merged_CDEs_9.18.20.csv"))

merged <- merged %>% 
  filter(USI != "Unknown") %>% 
  as.data.frame() %>%
  set_rownames(.$USI)

dim(merged)
```

```{r}
IDmap <- read.csv("References/St.Jude_TARGET_CICERO_Sample_IDmap.csv")

head(IDmap)
```
 
 
# Read in File Manifests 

```{r}
polyA_RNAseq_files <- read.csv("BAM_Manifests/TARGET_AML_polyA_RNAseq_Bam_Manifest_10.02.20.csv", 
                               row.names = 1) 


dim(polyA_RNAseq_files)
table(duplicated(polyA_RNAseq_files$Sample)) #FALSE
```

```{r}
RBD_RNAseq_files <- read.csv("BAM_Manifests/TARGET_AML_Ribodepleted_RNAseq_Bam_Manifest_10.02.20.csv",row.names = 1)

dim(RBD_RNAseq_files)
table(duplicated(RBD_RNAseq_files$Sample)) #FALSE
```

```{r}
SJ_RNAseq_files <- read.csv("BAM_Manifests/St.Jude_AML_ALL_RNAseq_Bam_Manifest_10.09.20.csv")

dim(SJ_RNAseq_files)
head(SJ_RNAseq_files)
```


# Read in the Contig Sequences

```{r message=FALSE, warning=FALSE}
TA.contigs <- read.csv("Data/TransAbyss_Contig_Sequences.csv")

# head(TA.contigs)
dim(TA.contigs) #54,878     4
```

```{r message=FALSE}
cicero.contig <- read.csv("Data/CICERO_Contig_Sequences.csv")

# head(cicero.contig)
dim(cicero.contig) #158,972 
```


# Read in the Theoretical Junc Sequences

```{r}
theoretical_data.collapsed <- read.csv("Data/Theoretical_Fusion_Breakpoint_Junctions_per_Transcript.csv")

head(theoretical_data.collapsed)
dim(theoretical_data.collapsed) # 1955    4
``` 


# Chromosome Lengths

```{r}
seqlens <- read.delim("Data/Grch37.lite_chrom_lengths.txt") %>% 
  rownames_to_column("chr") %>%
  pull(x, name=chr)

head(seqlens)
```

```{r}
seqlens.ctat <- read.delim("Data/Grch37.ctat_chrom_lengths.txt") %>% 
  rownames_to_column("chr") %>%
  pull(x, name=chr)

head(seqlens.ctat)
```


# Define Fusion Groups 

```{r}
Groups <- unique(theoretical_data.collapsed$Fusion.Category)

Groups
```


# Define defaults

I just need a breakpoint, but since I'll be searching for evidence in the RNAseq reads for the whole chromosome, don't need to exact breakpoint junction per patient. Just need a common one, such as 

```{r}
common_breakpoints <- read.csv("Data/TARGET_AML_Most_Common_RNAseq_Breakpoints.csv", 
                               row.names = 1) 


common_breakpoints 
```



# Run the Function 

## Rslurm 

```{r}
outdir <- file.path(SCRATCH,'jlsmith3/Fusion_Breakpoints')
sopt <- list(nodes='1', 'cpus-per-task'='2',
             'partition'='campus-new', 'mem'='40G',
             'time' = '48:00:00', 'mail-type'='FAIL,END',
             'mail-user'='jlsmith3@fredhutch.org') 
```


### polyA

```{r eval=FALSE}
setwd(outdir)
jobs1 <- lapply(Groups, 
       submit_slurm_per_fusion, 
       file_manifest=polyA_RNAseq_files,
       chromLengths=seqlens,
       cohort="polyA_RNAseq")
```

### RBD

```{r eval=FALSE}
setwd(outdir)
RBD_RNAseq_files <- RBD_RNAseq_files %>% 
  filter(grepl("AML|NBM|CD34_PB", Group), 
         grepl("diagnostic|NBM|CD34_NBM", Time_point))

jobs2 <- lapply(Groups, 
       submit_slurm_per_fusion, 
       file_manifest=RBD_RNAseq_files,
       chromLengths=seqlens,
       cohort="RBD_RNAseq")

```


### St.Jude

```{r eval=FALSE}
setwd(outdir)
jobs3 <- lapply(Groups, 
       submit_slurm_per_fusion, 
       file_manifest=SJ_RNAseq_files,
       chromLengths=seqlens.ctat,
       cohort="SJ_RNAseq")
```



## Furrr (Future map)

```{r eval=FALSE}
library(furrr)
future::plan("multisession") 
# future::plan("sequential") 
plan()
availableCores()
```

```{r eval=FALSE}
setwd(outdir)
for ( i in 1:length(Groups) ){ 
  # The fusion group to investigate
  group <- Groups[i]
  breakpoint <- common_breakpoints[group,"breakpoint"]
  print(group)
  print(breakpoint)
  
  #Filter the sequence data for the fusion to investigate
  fusion_theoretical_input <- filter(theoretical_data.collapsed, Fusion.Category == group)
  cicero.contig_input <- filter(cicero.contig, Fusion.Category ==  group)
  TA.contigs_input <- filter(TA.contigs, Fusion.Category == group)
  
  # I forgot seqs need to be the same width for pdict objects. hopefully will change soon.
  try(
    mut20 <- create_custom_pdict(theoretical_seqs_df = filter(fusion_theoretical_input,
                                                              grepl("Mutseq20", Name))), silent=T)
  try(
    mut25 <- create_custom_pdict(theoretical_seqs_df = filter(fusion_theoretical_input,
                                                            !grepl("Mutseq20", Name))), silent = T)
  
  for (n in 1:2){
  
    mutseq <- c("mut20","mut25")[n]
    try(pdict_obj <- get(mutseq), silent=T)
    fname <- paste0("Rdata/TARGET_AML_",group,"_polyA_RNAseq_",mutseq,".RDS")
    
    if(exists("pdict_obj")){ #some groups do not have both types of data.
      samps <- as.list(pull(polyA_RNAseq_files, Sample))
      tictoc::tic()
      polyA_evidence <- furrr::future_map_dfr(samps,
                            examine_breakpoint_evidence, 
                            breakpoint = breakpoint,
                            file_manifest = polyA_RNAseq_files,
                            Seqlens = seqlens, 
                            fusion_theoretical_pdict = pdict_obj, 
                            cicero.contig = cicero.contig_input, 
                            TA.contigs = TA.contigs_input,
                            .progress=TRUE)
      tictoc::toc()
      
      #Save output to RDS data
      saveRDS(polyA_evidence, fname)
      
      #remove large objects from memory
      rm(pdict_obj)
      rm(polyA_evidence) 
      gc()
    }
  }
}
```



# Session Information

```{r}
sessionInfo()
```



